---
pagetitle: "Share of the population living in extreme poverty"
output: 
  html_document:
    theme: lumen
    css: ["./assemblers/dsr.css"]
    df_print: paged
    mathjax: NULL
    code_folding: hide
    include:
      in_header: "./assemblers/header.html"
      after_body: "./assemblers/footer.html"
    #self_contained: false
    #lib_dir: libs
---

```{r setup, include=FALSE}
library(tidyverse)
library(tidyr)

df <- read.csv("data/share-of-the-population-living-in-extreme-poverty.csv")
df <- df %>% filter(Year == '2015') %>% rename("Ratio" = "Poverty.headcount.ratio.at..1.90.a.day..2011.PPP.....of.population.")

```

The above dataset includes the 

### SDG INDICATOR 1.1.1 - Share of the population living in extreme poverty, by country, in 2015:
The most common measures of <a href="https://en.wikipedia.org/wiki/Central_tendency">Central Tendency</a> are the mean, median and mode. Let's put these measures to practice through an example.


```{r, echo=FALSE}

df

```


#### Let's calculate this time series 

```{r, echo=FALSE}

mean <- round(mean(df$Ratio), digits = 2)
median <- median(df$Ratio)
mode <- df %>% count(Ratio) %>% arrange(n) %>% tail(1) %>% select(Ratio)
mode <- as.numeric(mode)

print(paste0("mean: ",mean))
print(paste0("mean: ",median))
print(paste0("mode: ",mode))

```

These measures seem far apart.

# let's plot a histogram to see what's going on

```{r, message=FALSE, echo=FALSE, fig.width=10.5, fig.height=7.5}

# "ARIMA"="#3189a3", "ETS"="#24bbed", "EIU"="#f26622")

ggplot(df, aes(Ratio)) +
  geom_histogram(bins = 100)+ 
  geom_vline(xintercept = mean, color = "#3189a3")+ 
  geom_text(aes( x=mean+2, y=20, label="mean"), color = "#3189a3") +
  geom_vline(xintercept = median, color = "#24bbed") +
  geom_text(aes( x=median+2, y=22, label="nedian"), color = "#24bbed") +
  geom_vline(xintercept = mode, color = "#f26622") + 
  geom_text(aes( x=mode+2, y=24, label="mode"), color = "#f26622") +
  theme(panel.grid = element_blank(), 
        #axis.ticks = element_blank(), 
        #axis.text.x=element_blank(), 
        #panel.border = element_blank(),
        strip.background = element_blank(),
        panel.background = element_blank()
        #axis.title.x=element_blank(),
        #axis.title.y=element_blank()
  ) 

```

<p>Median is mid-point where 50% of data is "behind" and "after", where Mean is the sum of values divided by number of elements. As this graph exemplify, mean is sensitive to outliers (here we may consider outliers the countries with Ratio above 10, which represents 15 countries out of 81 observations</p>

