---
pagetitle: "Has temperature risen in the past 100 years?"
output: 
  html_document:
    theme: lumen
    css: ["./assemblers/dsr.css"]
    df_print: paged
    mathjax: NULL
    code_folding: hide
    include:
      in_header: "./assemblers/header.html"
      after_body: "./assemblers/footer.html"
    #self_contained: false
    #lib_dir: libs
---
## Has temperature really risen in the past 100 years?
In this report, we put Time Series analysis into practice. We choose a simple measure, the monthly average temperature (In fahrenheit) measured in Central Park, New York from 1895 to today to 

```{r include=FALSE}
library(forecast)
library(ggplot2)
library(readr)
require(scales)
library(zoo)
library(naniar)
library(dplyr)
```

## Load data
We will be using the average monthly temperature measured in New York Central Park since 1895
https://www.ncdc.noaa.gov/cag/city/time-series/USW00094728

```{r }
# Load data
setwd("C:/RProjects/Kamino_R")
Central_Park_Temperature = read.csv("data/USW00094728-tavg-all-8-1895-2020.csv", header=TRUE, stringsAsFactors=FALSE, skip = 4)
head(Central_Park_Temperature)
```

## Data Cleanup


```{r, echo=FALSE}
# libraries: requires gglot2 and scales
# Notice one record that registers at Value = -99. This indicates a missing element for that specific date
# the next effort will handle this problem

ggplot(data=Central_Park_Temperature, aes(x=Date, y=Value, group=1)) +
  geom_line() +
  geom_point() + 
  scale_y_continuous(labels = comma) + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

We notice one odd measure (temperature of -99, which indicate abasence of data). The following code replace this data point with the value that came just before it.

```{r}
# Needs libraries naniar and dplyr
# replaces "-99" record with NA
CP_Temp <- Central_Park_Temperature %>% replace_with_na_all(condition = ~.x == -99)

# Needs library zoo
# Replaces NA records by the Value of the record that comes just before it
CP_Temp <- na.locf(CP_Temp)

# Redrawind the plot, all records are now within the expected band
ggplot(data=CP_Temp, aes(x=Date, y=Value, group=1)) +
  geom_line() +
  geom_point() + 
  scale_y_continuous(labels = comma) + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

```

## Can we answer the question already using a regression line?

We can already notice the trend upwards.

```{r}
# Adding a linear regression line, we already see the trend upwards. 
ggplot(data=CP_Temp, aes(x=Date, y=Value, group=1)) +
  geom_line() +
  geom_point() + 
  scale_y_continuous(labels = comma) + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  geom_smooth(method="lm", se=FALSE)

```

## Let's analyse the ACF, PACF, residual graphs

```{r}
# Time to create a univariate time-series
univar <- data.frame(CP_Temp$Value)

# Creating a time series
TimeSer <- ts(univar, start=c(1895,1), end=c(2020,8), frequency=12)

#acf(TimeSer)
ggAcf(TimeSer)

pacf(TimeSer)

plot(diff(TimeSer))

```

## Further useful visuals

```{r}
# Library: requires forecast
# autoplot(TimeSer) +
#   ggtitle("TimeSer") +
#   xlab("Month") +
#   ylab("Temperature (F)")

# Data exploration using graphs
ggseasonplot(TimeSer, year.labels = TRUE) +
  ylab("Temperature (F)") +
  ggtitle("Seasonal plot: Temperature per month")

ggseasonplot(TimeSer, year.labels = TRUE, polar=TRUE)+
  ylab("Temperature (F)") +
  ggtitle("Polar seasonal plot: Temperature per month")

# The horizontal lines indicate the means for each month
ggsubseriesplot(TimeSer)+
  ylab("Temperature (F)") +
  ggtitle("Seasonal subseries plot: Temperature per month")

# Analyse correlation on lags
lags <- window(TimeSer, start=1992)
gglagplot(lags)

# Analyse subset
lastFiveYears <- window(TimeSer, start=2015)
#autoplot(lastFiveYears) + xlab("Month") + ylab("Temperature (F)")

```

Let's "forecast" using naive assumptions: Mean, Naive, and Seasonal Naive predictions

```{r}
#The slow decrease in the ACF as the lags increase is due to the trend, 
# while the "scalloped" shape is due the seasonality.
ggAcf(lastFiveYears, lag=48)

#training <- window(TimeSer,start=2015,end=c(2019,1))
fit1 <- meanf(window(TimeSer, start=2015,end=c(2019,1)),h=24)
fit2 <- rwf(window(TimeSer, start=2015,end=c(2019,1)),h=24)
fit3 <- snaive(window(TimeSer, start=2015,end=c(2019,1)),h=24)
autoplot(window(TimeSer, start=2015)) +
  autolayer(fit1, series="Mean", PI=FALSE) +
  autolayer(fit2, series="Naïve", PI=FALSE) +
  autolayer(fit3, series="Seasonal naïve", PI=FALSE) +
  xlab("Year") + ylab("Temp") +
  ggtitle("Forecasts") +
  guides(colour=guide_legend(title="Forecast"))

# Decomposition
TimeSer %>% decompose(type="multiplicative") %>%
  autoplot() + xlab("Year") +
  ggtitle("Classical multiplicative decomposition
    of electrical equipment index")

#gghistogram(res) + ggtitle("Histogram of residuals")

#ggAcf(res) + ggtitle("ACF of residuals")

```

Time to run Forecasting Models:

```{r}
############### Forecasting Models ####################


####### Automated Forecasting using an ARIMA model

#window(TimeSer, start=2015,end=c(2019,1)) %>% forecast() %>% autoplot()
#window(TimeSer, start=2015,end=c(2019,1)) %>% ets() %>% autoplot()

#msBest <- auto.arima(TimeSer)
#msBest

#theForecast <- forecast(object=msBest, h=12)
#theForecast %>% autoplot()

# Automated Forecasting using an ARIMA model
fit_arima <- auto.arima(TimeSer)
fit_arima

forecast_arima <- forecast(object=fit_arima, h=24)
forecast_arima

#Plot forecast
autoplot(window(TimeSer, start=2015)) +
  autolayer(forecast_arima, series="ARIMA", PI=FALSE) +
  xlab("Year") + ylab("Temperature") +
  ggtitle("Forecast") +
  guides(colour=guide_legend(title="Forecast"))

#Including confidence intervals
autoplot(window(TimeSer, start=2015)) +
  autolayer(forecast_arima, series="ARIMA", PI=TRUE) +
  xlab("Year") + ylab("Temperature") +
  ggtitle("Forecast") +
  guides(colour=guide_legend(title="Forecast"))

#autoplot(window(forecast_arima, start=2015))

```

```{r}
###### Exponential Smoothing

fit_ets <- ets(TimeSer, model="ZZZ")
summary(fit_ets)

forecast_ets <- forecast(object=fit_ets, h=24)
forecast_ets

autoplot(window(TimeSer, start=2015)) +
  autolayer(forecast_ets, series="ETS", PI=TRUE) +
  xlab("Year") + ylab("Temperature") +
  ggtitle("Forecast") +
  guides(colour=guide_legend(title="Forecast"))

```


