---
pagetitle: "Web scrapping - Find sunniest cities in the World"
output: 
  html_document:
    theme: flatly
    code_folding: show
    df_print: paged
---

```{r setup, include=FALSE}
library(rvest)
library(tidyverse)
library(tidyr)
library(DT)
library(reactable)

```


### What are the sunniest cities in the world, by continent:


<p>The data we need in stored in several tables on this <a href="https://en.wikipedia.org/wiki/List_of_cities_by_sunshine_duration">Wikipedia page - List of cities by sunshine duration</a>. The data records the sunshine hours by month, and by year. Let's use web scrapping to capture this information</p>

### First, let's grab the table for Africa

```{r, echo=FALSE}

url = 'https://en.wikipedia.org/wiki/List_of_cities_by_sunshine_duration'

africa = url %>% 
    read_html() %>%
    html_node(xpath = '//*[@id="mw-content-text"]/div[1]/table[2]') %>%
    html_table(fill = TRUE)
africa

```


```{r, echo=FALSE}

# Grabbing the data for the other continents
asia = url %>% 
    read_html() %>%
    html_node(xpath = '//*[@id="mw-content-text"]/div[1]/table[3]') %>%
    html_table(fill = TRUE)

europe = url %>% 
  read_html() %>%
  html_node(xpath = '//*[@id="mw-content-text"]/div[1]/table[4]') %>%
  html_table(fill = TRUE)

north_central_america = url %>% 
  read_html() %>%
  html_node(xpath = '//*[@id="mw-content-text"]/div[1]/table[5]') %>%
  html_table(fill = TRUE)

south_america = url %>% 
  read_html() %>%
  html_node(xpath = '//*[@id="mw-content-text"]/div[1]/table[6]') %>%
  html_table(fill = TRUE)

america <- rbind(north_central_america, south_america)

oceania = url %>% 
  read_html() %>%
  html_node(xpath = '//*[@id="mw-content-text"]/div[1]/table[7]') %>%
  html_table(fill = TRUE)

tabledata <- rbind(africa, asia, europe, america, oceania) %>% mutate(Year = round(as.numeric(gsub(",", "", Year)),digits = 0))

```

### Let's now grab the data for all continents, and turn it into a searchable table

```{r table, echo=FALSE, message=FALSE, warning=FALSE}
#DT::datatable(tabledata, filter = 'top')
rearranged <- tabledata[,c(1,2,15,3:14)]

reactable(rearranged, filterable = TRUE, columns = list(Jan = colDef(filterable = FALSE), Feb = colDef(filterable = FALSE), Mar = colDef(filterable = FALSE),
                                                        Apr = colDef(filterable = FALSE),May = colDef(filterable = FALSE),Jun = colDef(filterable = FALSE),
                                                        Jul = colDef(filterable = FALSE),Aug = colDef(filterable = FALSE),Sep = colDef(filterable = FALSE),
                                                        Oct = colDef(filterable = FALSE),Nov = colDef(filterable = FALSE),Dec = colDef(filterable = FALSE),
                                                        Year = colDef(filterable = FALSE)))
# reactable(
#   tabledata,
#   searchable = TRUE,
#   details = function(index) {
#     levels <- filter(entity_levels, ENTITY == entity_levels_grp$ENTITY[index]) %>% select(-ENTITY)
#     tbl <- reactable(levels, outlined = TRUE, highlight = TRUE, fullWidth = TRUE)
#     htmltools::div(style = list(margin = "10px 10px"), tbl)
#   },
#   onClick = "expand",
#   rowStyle = list(cursor = "pointer" )
# )

```

### Time to plot the top 10 sunniest cities on each contient


```{r, echo=FALSE}

# Preparing the data for the visual
to_add <- data.frame(
  individual=c(NA,NA,NA),
  value=c(NA,NA,NA))

top_10_africa <- africa %>%
  mutate(num_year = as.numeric(gsub(",", "", Year))) %>% arrange(desc(num_year)) %>% 
  head(10) %>% 
  select(City, num_year) %>% rename(individual = City, value = num_year)

top_10_africa <- rbind(top_10_africa, to_add)%>% mutate(group = "africa")

top_10_asia <- asia %>%
  mutate(num_year = as.numeric(gsub(",", "", Year))) %>% arrange(desc(num_year)) %>% 
  head(10) %>% 
  select(City, num_year) %>% rename(individual = City, value = num_year)

top_10_asia <- rbind(top_10_asia, to_add) %>% mutate(group = "asia")

top_10_europe <- europe %>%
  mutate(num_year = as.numeric(gsub(",", "", Year))) %>% arrange(desc(num_year)) %>% 
  head(10) %>% 
  select(City, num_year) %>% rename(individual = City, value = num_year)

top_10_europe <- rbind(top_10_europe, to_add) %>% mutate(group = "europe")

top_10_america <- america %>%
  mutate(num_year = as.numeric(gsub(",", "", Year))) %>% arrange(desc(num_year)) %>% 
  head(10) %>% 
  select(City, num_year) %>% rename(individual = City, value = num_year)

top_10_america <- rbind(top_10_america, to_add) %>% mutate(group = "america")

top_10_oceania <- oceania %>%
  mutate(num_year = as.numeric(gsub(",", "", Year))) %>% arrange(desc(num_year)) %>% 
  head(10) %>% 
  select(City, num_year) %>% rename(individual = City, value = num_year)

top_10_oceania <- rbind(top_10_oceania, to_add) %>% mutate(group = "oceania")

data <- rbind(top_10_africa, top_10_asia, top_10_europe, top_10_america, top_10_oceania) 

```



```{r echo=FALSE, fig.height=7.5, fig.width=10.5, message=FALSE, warning=FALSE}

empty_bar <- 3

data <- data %>% arrange(group)
data$id <- seq(1, nrow(data))

label_data <- data
number_of_bar <- nrow(label_data)
angle <- 90 - 360 * (label_data$id-0.5) /number_of_bar     # I substract 0.5 because the letter must have the angle of the center of the bars. Not extreme right(1) or extreme left (0)
label_data$hjust <- ifelse( angle < -90, 1, 0)
label_data$angle <- ifelse(angle < -90, angle+180, angle)

# prepare a data frame for base lines
base_data <- data %>% 
  group_by(group) %>% 
  summarize(start=min(id), end=max(id) - empty_bar) %>% 
  rowwise() %>% 
  mutate(title=mean(c(start, end)))

grid_data <- base_data
grid_data$end <- grid_data$end[ c( nrow(grid_data), 1:nrow(grid_data)-1)] + 1
grid_data$start <- grid_data$start - 1
grid_data <- grid_data[-1,]

# Order data:
data = data %>% arrange(id)

v1 <- ggplot(data, aes(x=as.factor(id), y=value, fill=group)) +       # Note that id is a factor. If x is numeric, there is some space between the first bar
  
  geom_bar(aes(x=as.factor(id), y=value, fill=group), stat="identity", alpha=0.5) +
  
  # Add a val=100/75/50/25 lines. I do it at the beginning to make sur barplots are OVER it.
#  geom_segment(data=grid_data, aes(x = 0, y = 5000, xend = 44, yend = 5000), colour = "grey", alpha=1, size=0.3 , inherit.aes = FALSE ) +
  geom_segment(data=grid_data, aes(x = 0, y = 4000, xend = 44, yend = 4000), colour = "grey", alpha=0.1, size=0.3 , inherit.aes = FALSE ) +
  geom_segment(data=grid_data, aes(x = 0, y = 3000, xend = 44, yend = 3000), colour = "grey", alpha=0.1, size=0.1 , inherit.aes = FALSE ) +
  geom_segment(data=grid_data, aes(x = 0, y = 2000, xend = 44, yend = 2000), colour = "grey", alpha=0.1, size=0.1 , inherit.aes = FALSE ) +
  
  # Add text showing the value of each 100/75/50/25 lines
  annotate("text", x = rep(max(data$id),3), y = c(2000, 3000, 4000), label = c("2000 hours", "3000 hours", "4000 hours") , color="grey", size=3 , angle=0, fontface="bold", hjust=1) +
  
  geom_bar(aes(x=as.factor(id), y=value, fill=group), stat="identity", alpha=0.5) +
  ylim(-2500,5000) +
  theme_minimal() +
  theme(
    legend.position = "none",
    axis.text = element_blank(),
    axis.title = element_blank(),
    panel.grid = element_blank(),
    plot.margin = unit(rep(-1,4), "cm") 
  ) +
  coord_polar() + 
  # Add labels on top of each bar
  geom_text(data=label_data, aes(x=id, y=value+100, label=paste0(individual,": ",round(value,digits=0)), hjust=hjust), color="black", fontface="bold",alpha=0.6, size=4, angle= label_data$angle, inherit.aes = FALSE ) +
  
  # Add base line information
#  geom_segment(data=base_data, aes(x = start, y = -2000, xend = end, yend = -2000), colour = "black", alpha=0.8, size=0.6 , inherit.aes = FALSE )  +
  geom_text(data=base_data, aes(x = title, y = -500, label=group), colour = "black", alpha=0.8, size=4, fontface="bold", inherit.aes = FALSE)

v1

```

